Interoception Scale - Data Cleaning

### 2024-11-30 study1/analysis/1_cleaning.qmd
update

library(tidyverse)
library(easystats)
library(patchwork)
library(ggside)

@@

p_feed <- estimate_density(df$Experiment_Enjoyment) |>
  normalize(select = y) |>
  mutate(y = y * 86) |>  # To match the binwidth
  ggplot(aes(x = x)) +
  geom_bar(data=df, aes(x = Experiment_Enjoyment, fill = Condition), bins=28) +
  scale_fill_manual(values = c("Dimensions"= "#ff944d","Domains" = "#ca0020", "Random"="#ff80d5")) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Experiment Enjoyment and Condition", y = "Number of Participants", fill = " ", x = "Ratings of enjoyment on a scale 1-5", subtitle = "Distribution of Participants' Experiment Enjoyment and Condition") +
  facet_wrap(~Condition) +
  theme_modern(axis.title.space = 10) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(1.1)),
  )
p_feed

@@

```{r}
data.frame(Condition = df$Condition,
          Comments = df$Experiment_Feedback) |> 
  filter(!tolower(Comments) %in% c(NA, "None", "no", "none", "No thank you", "no thank you", "no not at all", "no thanks", "No comment", "No", "nil", "N.A", "n/a", "N/A", "na", "Not Really")) |> 
  gt::gt() |> 
  gt::opt_stylize() |> 
  gt::opt_interactive(use_compact_mode = TRUE) 
```

@@

```{r}
#| warning: false
#| error: false

dfchecks <- df |>
  dplyr::mutate(
    # "I always know that I am attentively doing a study"
    A1 = ifelse(Sexual_State_A == 6 | Sexual_State_A == 5, 0, 1),
    # "Even if I am anxious, I should now answer all the way to the left"
    A2 = ifelse(Anxious_SkinThermo_A == 0, 0, 1),
    # "I often experience sensations, and I will answer zero to this question"
    A3 = ifelse(Nociception_ColonBladder_A == 0, 0, 1),
    # "In general, I am very sensitive and attentive to the questions I am currently answering"
    A4 = ifelse(Sensitivity_Cardiac_A == 6 | Sensitivity_Cardiac_A == 5, 0, 1),
    # "I often pay attention to the answers I am giving"
    A5 = ifelse(Sensitivity_Gastric_A == 6 | Sensitivity_Gastric_A == 5, 0, 1),
    # "I can always accurately answer to the left on this question to show that I am reading it"
    A6 = ifelse(Accuracy_Respiratory_A == 0 | Accuracy_Respiratory_A == 1, 0, 1),
    # "I can always accurately perceive that to this question I should answer the lowest option"
    A7 = ifelse(Accuracy_Genital_A == 0, 0, 1),
    # "Sometimes I notice that I need to answer all the way to the right"
    A8 = ifelse(Confusion_ColonBladder_A == 6, 0, 1),
    .keep = "none"
  ) 
dfchecks$Total <- rowSums(dfchecks)

@@

### Mahalanobis Distance

@@

# parameters::parameters(glm(Outlier_AttentionCheck ~ Distance_LOF, data = dfoutlier, family = "binomial"))

@@

This measure takes into account correlations within the dataset, which is especially useful when trying to refine and validate a questionnaire with overlapping itmes.

@@

```{r}
df <- filter(df, !Participant %in% outliers$attentionchecks)
df <- filter(df, !Participant %in% outliers$distance)
```

>>>>

library(tidyverse)
library(easystats)
library(patchwork)
library(ggside)
library(ggdist)

@@

```{r}
df |> 
  summarise(n = n(), .by=c("Condition", "Experiment_Enjoyment")) |>
  filter(!is.na(Experiment_Enjoyment)) |> 
  mutate(n = n / sum(n),
         Experiment_Enjoyment = fct_rev(as.factor(Experiment_Enjoyment)),
         .by="Condition") |>
  ggplot(aes(y = Condition, x = n, fill = Experiment_Enjoyment)) +
  geom_bar(stat="identity", position="stack") +
  scale_fill_manual(values=c("green", "yellowgreen",  "yellow", "orange", "red")) +
  scale_x_continuous(expand=c(0, 0))  +
  labs(title="Experiment Enjoyment",
       subtitle="Proportion of participants by condition") + 
  guides(fill = guide_legend(reverse=TRUE)) +
  theme_minimal() +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        legend.position = "top",
        legend.title = element_blank())

lm(Experiment_Enjoyment ~ Condition, data = df) |> 
  modelbased::estimate_contrasts(p_adjust="none", contrast = "Condition") |> 
  display()
```

@@

```{r}
data.frame(Comments = df$Experiment_Feedback) |> 
  filter(!tolower(Comments) %in% c(NA, "no", "none", "no thank you", "no not at all", "no thanks", "no comment", "nil", "n.a", "n/a", "na", "not really", "thank you", "nothing", "no not at the moment", "none for now", "thank you!", "thanks", "not really.")) |> 
  gt::gt() |> 
  gt::opt_stylize() |> 
  gt::opt_interactive(use_compact_mode = TRUE) 
```

@@

```{r}
dfchecks <- df |>
  dplyr::mutate(
    # "I always know that I am attentively doing a study"
    A1 = ifelse(Sexual_State_A == 6 | Sexual_State_A == 5, 0, 1),
    # "Even if I am anxious, I should now answer all the way to the left"
    A2 = ifelse(Anxious_SkinThermo_A == 0, 0, 1),
    # "I often experience sensations, and I will answer zero to this question"
    A3 = ifelse(Nociception_ColonBladder_A == 0, 0, 1),
    # "In general, I am very sensitive and attentive to the questions I am currently answering"
    A4 = ifelse(Sensitivity_Cardiac_A == 6 | Sensitivity_Cardiac_A == 5, 0, 1),
    # "I often pay attention to the answers I am giving"
    A5 = ifelse(Sensitivity_Gastric_A == 6 | Sensitivity_Gastric_A == 5, 0, 1),
    # "I can always accurately answer to the left on this question to show that I am reading it"
    A6 = ifelse(Accuracy_Respiratory_A == 0 | Accuracy_Respiratory_A == 1, 0, 1),
    # "I can always accurately perceive that to this question I should answer the lowest option"
    A7 = ifelse(Accuracy_Genital_A == 0, 0, 1),
    # "Sometimes I notice that I need to answer all the way to the right"
    A8 = ifelse(Confusion_ColonBladder_A == 6, 0, 1),
    .keep = "none"
  ) 
dfchecks$Total <- rowSums(dfchecks)

@@

### Multivariate Distance

@@

@@

This measure takes into account correlations within the dataset, which is especially useful when trying to refine and validate a questionnaire with overlapping items.

@@

### Predictors

```{r}
dfoutlier <- full_join(dfoutlier, select(df, Participant, Experiment_Duration), by = join_by(Participant))

parameters::parameters(glm(Outlier_AttentionCheck ~ Distance_LOF, data = dfoutlier, family = "binomial"))
parameters::parameters(glm(Outlier_AttentionCheck ~ poly(Experiment_Duration, 2), data = dfoutlier, family = "binomial"))
parameters::parameters(glm(Distance_LOF ~ poly(Experiment_Duration, 2), data = dfoutlier))
```

```{r}
r_full <- cor(select(df, contains("_Q")))
r_cleaned <- filter(df, !Participant %in% c(outliers$attentionchecks, outliers$distance)) |> 
  select(contains("_Q")) |> 
  cor()

data.frame(Full=r_full[lower.tri(r_full)], Clean=r_cleaned[lower.tri(r_cleaned)]) |> 
  pivot_longer(everything()) |> 
  ggplot(aes(x=value)) +
  geom_histogram(aes(fill=name, color=name), bins=60, alpha=0.3, position="identity") +
  theme_minimal()
```

```{r}
#| code-fold: false

df <- filter(df, !Participant %in% c(outliers$attentionchecks, outliers$distance))
```
